---
title: "billboard 100"
author: "kat"
date: "9/15/2021"
output: html_document
---

## foreword + some loading in

This is actually a dataset I'm -very- familiar with, as I'm working on a pub with it right now. However, we manually pulled all this data from the internet and uisng Spotify's API, so maybe I should have foreseen this tidy tuesday!

```{r}
#lil load ins
library(tidyverse)
library(tidytuesdayR)
library(futurevisions)
library(extrafont)
library(lubridate)
library(sysfonts)
font_add_google("Source Sans Pro")
font_add_google("Josefin Sans")
font_add_google("Bebas Neue")

#load the data in
tuesdata <- tidytuesdayR::tt_load('2021-09-14')

#let's separate the two
billboard <- tuesdata$billboard #billboard first

features <- tuesdata$audio_features #audio features second
``` 

## lil peekaboo

I want to take a look first. I see that we have the spotify api data in the "audio features" csv and the billboard stats in the "billboard.csv" file. 

Below is the 'billboard' file
```{r}
glimpse(billboard) # ok, so we can see our variables
```

And now the 'features' file.
```{r}
glimpse(features)
```

## a confession

Maybe we can do a top perfomers (denoted by both week_position being #1 and "weeks_on_chart") for the decades! We'll start with 70s.

Unfortunately, dates (like, days -and- the events) scare me, even with lubridate().

```{r}
billboard$week_id <- mdy(billboard$week_id)
```

## Open up the dates pit
70s
```{r}
billboard_70s <- billboard %>%
  filter(lubridate::year(week_id) %in% c(1970:1979))
```

80s
```{r}
billboard_80s <- billboard %>%
  filter(lubridate::year(week_id) %in% c(1980:1989))
```

90s
```{r}
billboard_90s <- billboard %>%
  filter(lubridate::year(week_id) %in% c(1990:1999))
```

2000s
```{r}
billboard_2000s <- billboard %>%
  filter(lubridate::year(week_id) %in% c(2000:2009))
```

2010s
```{r}
billboard_2010s <- billboard %>%
  filter(lubridate::year(week_id) %in% c(2010:2019))
```

Let's find the top 10 number 1 songs for the 70s
```{r}
seventies <- billboard_70s %>%
  group_by(song, performer) %>%
  filter(peak_position == 1) %>%
  filter(peak_position == week_position)%>%
  count(peak_position, sort = TRUE) %>%
  head(seventies, n = 10)

#create new decade column and populate it with 'seventies'
seventies$decade <- c('70s') 
```


Let's find the top 10 number 1 songs for the 80s
```{r}
eighties <- billboard_80s %>%
  group_by(song, performer) %>%
  filter(peak_position == 1) %>%
  filter(peak_position == week_position)%>%
  count(peak_position, sort = TRUE) %>%
  head(eighties, n = 10)

#create new decade column and populate it with 'eighties'
eighties$decade <- c('80s') 
```

Let's find the top 10 number 1 songs for the 90s
```{r}
nineties <- billboard_90s %>%
  group_by(song, performer) %>%
  filter(peak_position == 1) %>%
  filter(peak_position == week_position)%>%
  count(peak_position, sort = TRUE) %>%
  head(nineties, n = 10)

#create new decade column and populate it with 'nineties'
nineties$decade <- c('90s') 
```


Let's find the top 10 number 1 songs for the 2000s
```{r}
two_thousands <- billboard_2000s %>%
  group_by(song, performer) %>%
  filter(peak_position == 1) %>%
  filter(peak_position == week_position)%>%
  count(peak_position, sort = TRUE) %>%
  head(two_thousands, n = 10)

#create new decade column and populate it with 'nineties'
two_thousands$decade <- c('2000s') 
```

Let's find the top 10 number 1 songs for the 2010s
```{r}
two_thousand_tens <- billboard_2010s %>%
  group_by(song, performer) %>%
  filter(peak_position == 1) %>%
  filter(peak_position == week_position)%>%
  count(peak_position, sort = TRUE) %>%
  head(two_thousand_tens, n = 10)

#create new decade column and populate it with 'two_thousand_tens'
two_thousand_tens$decade <- c('2010s') 
```


Full set:
When many files love each other very much, the etl stork brings them a baby called 'top_all_time'.
```{r}
top_all_time <- rbind(seventies, eighties, nineties, two_thousands, two_thousand_tens)
```

I'm going to rename the one song with two names to not throw off my plot.
```{r}
top_all_time$song[22]="Candle In The Wind"
```

plot
```{r, fig.width = 10, fig.height = 8}
p <- ggplot(top_all_time, aes(x=reorder(song,n), n, fill = decade, color = decade)) +
  geom_bar(stat="identity") +
  coord_flip() + 
  theme_minimal()


final <- p + 
  annotate("text",  x = 11, y = 25, family = "Source Sans Pro", label = "Billboard\n100's\nTop 10\n#1 Songs\nby Decade", hjust = 1, vjust = 0, colour = "black", size = 18, lineheight = 1) +
  annotate("text",  x = 5, y = 25, family = "Lucida Sans", label = "The top 10 songs that\n stayed at Number 1\n from the 1970s-2010s", hjust = 1, vjust = 0, colour = "#09283C", size = 6, lineheight = 1) +
  labs(caption = "Tidy Tuesday | Week 38 | @kat__chia") +
  scale_color_manual(values = futurevisions("atomic_clock"), labels = c("70s", "80s", "90s", "2000s", "2010s")) +
  scale_fill_manual(values = futurevisions("atomic_clock"), labels = c("70s", "80s", "90s", "2000s", "2010s")) +
   theme(
    axis.title.x=element_blank(),
    axis.title.y=element_blank(),
    axis.text.y = element_text(family = "Lucida Sans"),
    legend.position = c(0.77, 0.05),
    legend.direction = "horizontal",
    legend.spacing.x = unit(0, "pt"),
    legend.key.width = unit(40, "pt"),
    legend.title = element_blank(),
    legend.text = element_text(family = "Source Sans Pro", size = 10),
    plot.background = element_rect(fill = "#f8f0e5", colour = NA),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank()
   ) +
  guides(color = guide_legend(label.position = "bottom")) 

print(final)
```

